---
title: "Turning the wheels of time --- A reproducibility case study"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=TRUE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Installing R and RStudio

* You can get R from CRAN: <https://cran.r-project.org/>
* Then, install RStudio: <https://www.rstudio.com/>
* You can install packages easily from within RStudio via  a call like this:

```{r, eval = FALSE}
install.packages("rmdformats")
```

* And then load them via a call like this:

```{r, eval = FALSE}
library("rmdformats")
```


# Our case study: Turning the wheels of time

[Topolinski and Sparenberg](ttp://dx.doi.org/10.1177/1948550611419266) (2012, press
release in German [here](https://www.uni-wuerzburg.de/sonstiges/meldungen/single/artikel/wer-mit-de/)) 
report that 
clockwise movements induce psychological states of temporal 
progression and an orientation toward the future and novelty.

Among other experiments, they tested whether self--reported experiential 
openness is influenced by rotational movements.

The authors use the "Openness to Experience" subscale from the
"Neuroticism -- Extroversion -- Openness" personality inventory ([NEO](https://en.wikipedia.org/wiki/Revised_NEO_Personality_Inventory#Personality_dimensions))
for their study. 

They found that participants who turn kitchen rolls clockwise 
report more openness to new experiences (Experiment 2 in their paper),
if one controls for mood and arousal.  

In 2015 [Wagenmakers et. al.](http://dx.doi.org/10.3389/fpsyg.2015.00494) 
published a replications study that could not
confirm these results.

The full data are available at <https://osf.io/uszvx/>, and as part of the JASP 
program with the name `Kitchen_Rolls.csv`. 

# Importing the data 

Here, we first load the R--package "[tidyverse](http://tidyverse.org/)". It acts
as a wrapper for a lot of very useful R packages. As the data comes in a
csv format, we can use the function `read_csv` from the tidyverse to import it
and display a preview.

Note that the function `read_csv` will print a message indicating the 
column specifications it has picked up.

The data can be conveniently viewed using `View()` in RStudio.

```{r import_data}
library(tidyverse)
k_rolls <- read_csv("../data_and_papers_reproducibility/Kitchen_Rolls.csv")
k_rolls
```

# Produce a beeswarm plot of the data 

We can check graphically, whether there is a relationship between
turning and openness by looking at the  **mean_Neo** variable. We use
the package **ggplot2** for the plotting.

```{r plot_data, dependson="import_data"}
d_for_ana <- select(k_rolls, ParticipantNumber, Rotation, mean_NEO) 

library(ggbeeswarm) 
library(Hmisc)


ggplot(d_for_ana, aes(x = Rotation, y = mean_NEO, color = Rotation)) +
      geom_beeswarm() +
      scale_color_hue(h.start = 30, l = 50) + 
      stat_summary(fun.data = mean_cl_normal, geom  = "crossbar")


```

The authors of the replication study found that the mean NEO score was significantly 
lower (albeit with a small effect size) in the clock--wise compared to the 
counter--clockwise group. 

Looking at the plot, we can actually observe that this, The boxes show the 
mean of the data and its 95% confidence interval. We can next try
to confirm the visual impression via a t-test.

# Perform a t-test to compare the two groups

We can now run a t-test to check whether our visual impression is correct, 
i.e. that there isn't actually a difference between the two experimental groups.

```{r perform_test, dependson="plot_data"}
t.test(mean_NEO ~ Rotation, data = d_for_ana)
```



